\begin{acknowledgements}
First of all, I want to thank both my advisors,
Prof. Matthias Hauswirth and Prof. Nate Nystrom.
Their expertise helped me to become a better researcher.
They taught me how to ask the proper questions when facing a problem,
that not only will help me in my professional career,
but in everyday life.

In our \unsafe{} study,
Luca Ponzanellli, Andrea Mocci and Prof. Michele Lanza
collaborated with us to correlate \unsafe{} 
comments in the Questions and Answers database, \stackoverflow{}, mentioned in Chapter~\ref{cha:unsafe}.
Thanks to them.

Special thanks to Max Sch\"afer from Semmle.
He executed our \ql{} queries in the entire \lgtm{} database used in Chapter~\ref{cha:casts}.
Without his help, this thesis would not have been possible.

Prof. Antonio Carzaniga and Prof. Gabriele Bavota made valuable comments during the Ph.D. proposal.
They contributed to find an appropriate sample size in the cast study.

My group colleagues, Mohammad, Dmitri, Andrea, and Igor.
With whom we had uncountable talks during my Ph.D.
Special mention for Amanj and Nosheen,
I learnt a lot in our coffee breaks.
Thank you guys.

On the personal side,
Vanesa, Cinzia, Mariano, Esteban, Dino, Olaf, Dimos, Jeff, Marisa, Rali, Gabi, Dilly, Sebastian, Eva, Stefania, and Grace 
who make me feel I found a family in Lugano.
My former and current flat mates, Nora, Marie-line, Lea, and Celine;
and Tanya and Matej, living with you guys was a pleasure.

My flat mate and dearest friend, Ankur Singhal,
I wish you were here to see this moment.
The Singhal family, Anil, Anita, and Ankita.
I appreciate your support in the toughest moments.

And last but not least,
I want to thank my family, for their unconditional support,
and for always believing in me.

\end{acknowledgements}
