\newcommand{\USI}{Universit\`a della Svizzera Italiana}

\newcommand{\cf}{\textit{cf.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\etc}{\textit{etc.}}
\newcommand{\adhoc}{\textit{ad-hoc}}
\newcommand{\perse}{\textit{per se}}
\newcommand{\circa}{\textit{circa}}

\newcommand{\phd}{Ph.D.}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\lang}[1]{\textsc{#1}}
\newcommand{\host}[1]{\textsl{#1}}
\newcommand{\conf}[2]{\textsf{#1'#2}}

\newcommand{\java}{\lang{Java}}
\newcommand{\csharp}{\lang{C\#}}
\newcommand{\cc}{\lang{C}}
\newcommand{\cpp}{\lang{C++}}
\newcommand{\smalltalk}{\lang{Smalltalk}}
\newcommand{\javascript}{\lang{JavaScript}}
\newcommand{\haskell}{\lang{Haskell}}
\newcommand{\scala}{\lang{Scala}}
\newcommand{\clojure}{\lang{Clojure}}
\newcommand{\ql}{\lang{QL}}
\newcommand{\fortran}{\lang{Fortran}}
\newcommand{\cobol}{\lang{Cobol}}
\newcommand{\pascal}{\lang{Pascal}}
\newcommand{\apl}{\lang{APL}}
\newcommand{\basic}{\lang{Basic}}
\newcommand{\swift}{\lang{Swift}}
\newcommand{\php}{\lang{PHP}}
\newcommand{\racket}{\lang{Racket}}
\newcommand{\rust}{\lang{Rust}}
\newcommand{\kotlin}{\lang{Kotlin}}
\newcommand{\ml}{\lang{SML}}

\newcommand{\api}{API}
\newcommand{\asm}{ASM}
\newcommand{\jnif}{JNIF}
\newcommand{\jvm}{JVM}
\newcommand{\jvmti}{JVMTI}
\newcommand{\jni}{JNI}
\newcommand{\lilith}{Lilith}

\newcommand{\smu}{\code{sun.misc.Unsafe}}

\newcommand{\github}{\host{GitHub}}
\newcommand{\gitlab}{\host{GitLab}}
\newcommand{\bitbucket}{\host{Bitbucket}}
\newcommand{\sourceforge}{\host{SourceForge}}
\newcommand{\maven}{\host{Maven}}
\newcommand{\mavencentral}{\host{Maven Central}}
\newcommand{\npm}{\host{npm}}
\newcommand{\boa}{\host{Boa}}
\newcommand{\candoia}{\host{Candoia}}
\newcommand{\lgtm}{\host{lgtm}}
\newcommand{\rascal}{\host{Rascal}}
\newcommand{\sourcegraph}{\host{sourcegraph}}
\newcommand{\bitly}{\host{Bitly}}

\newcommand{\hdr}{\rowcolor{header-color}}
\newcommand{\alt}{\rowcolor{alt-row-color}}
\newcommand{\row}{}

\newcommand{\unsafe}{\emph{Unsafe}}

\definecolor{header-color}{HTML}{D1D1D1}
\definecolor{alt-row-color}{HTML}{ECECEC}

\newcommand{\artstyle}[1]{\textsl{#1}}
\newcommand{\artexp}[2]{\artstyle{#1}:\artstyle{#2}}
\newcommand{\art}[3]{\artstyle{#1}:\artstyle{#2}\footnote{\url{#3}}}

\newcommand{\member}[1]{\emph{#1}}
\newcommand{\smugroup}[1]{\textsl{#1}}
\newcommand{\stackoverflow}{Stack Overflow}

\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\definecolor{lightgray}{rgb}{0.98,0.98,0.98}
\definecolor{lightblue}{rgb}{0.85,0.95,1}
\definecolor{gray}{rgb}{0.2,0.2,0.2}

% \definecolor{light-orange}{rgb}{1,0.6,0}

\definecolor{diffstart}{named}{Grey}
\definecolor{diffincl}{named}{Green}
\definecolor{diffrem}{named}{OrangeRed}
\definecolor{hl}{named}{OrangeRed}

\setminted{
	baselinestretch=1,
	bgcolor=lightgray,
	fontsize=\footnotesize,
	highlightcolor=lightblue,
	tabsize=4,
	linenos,
	escapeinside=\#\#
}

\newcommand{\urqA}{To what extent does the Unsafe API impact common application code?}
\newcommand{\urqB}{How and when are Unsafe features used?}

\newcommand{\urqAdesc}{We want to understand to what extent code actually uses Unsafe or depends on it.}
\newcommand{\urqBdesc}{We want to investigate what functionality third-party libraries require from Unsafe. This could point out ways in which the \java{} language and/or the JVM can be evolved to provide the same functionality, but in a safer way.}


\newcommand{\crqA}{How frequently is casting used in common application code?}
\newcommand{\crqB}{How and when casts are used?}
\newcommand{\crqC}{How recurrent are the patterns for which casts are used?}

\newcommand{\crqAdesc}{To what extent does application code actually use casting operations?}
\newcommand{\crqBdesc}{If casts are used in application code, how and when do developers use them?}
\newcommand{\crqCdesc}{In addition to understand how and when casts are used, we want to measure how often developers need to resort to certain idioms to solve a particular problem.}
