\begin{pattern}{Equals}
This pattern is a common pattern to implement the well-known \code{equals} method (declared in \code{java.lang.Object}).
It is a particularly instance of guarded casts.
A cast expression is guarded by either an
\code{instanceof} test---\variant{InstanceOf} variant---or a
\code{getClass} comparison---\variant{GetClass} variant---usually to the same target type as the cast;
in an \code{equals}%
\footnote{\url{https://docs.oracle.com/javase/8/docs/api/java/lang/Object.html\#equals-java.lang.Object-}}
method implementation.
This is done to check if the argument has same type as the receiver
(\code{this} argument).
Notice that a cast in an \code{equals} method is needed because it
receives an \code{Object} as a parameter.

To detect this pattern,
a cast must be applied to the parameter of the \code{equals} method.
The result value of the cast must be then used in an equality comparison.
We relax the constraint that the target type of the cast must the enclosing class.

\instances{}
This pattern accounts for \nEqualsOutOfGuarded\% of guarded casts,
\nEqualsPattern{} instances out of \nGuarded{}.
Figure~\ref{fig:casts:equals} shows the different variants of the \thisp{} pattern and their occurrences.
The \variant{InstanceOfSupertype}, \variant{AutoValue}, and
\variant{InstanceOfSwitch} variants are explained below.

\plot{patterns/table-pattern-Equals-Equals-args}{fig:casts:equals}{\thisp{} Variant Occurrences}

The following listing shows an example of the \thisp{} pattern.
In this case,
\code{instanceof} is used to guard for the same type as the receiver 
(\variant{InstanceOf} variant).

%https://lgtm.com/projects/g/neo4j/neo4j/snapshot/dist-15760049-1519892555006/files/community/kernel/src/main/java/org/neo4j/kernel/impl/api/CountsRecordState.java?sort=name&dir=ASC&mode=heatmap&excluded=false#L182
\def\urlvar{http://bit.ly/neo4j_neo4j_2vJw94J}
\begin{minted}[highlightlines=7]{java}
@Override
public boolean equals(Object obj) {
    if ( this == obj ) {
        return true;
    }
    if ( (obj instanceof Difference) ) {
        Difference that = (Difference) obj;
        return actualFirst == that.actualFirst
          && expectedFirst == that.expectedFirst
          && actualSecond == that.actualSecond 
          && expectedSecond == that.expectedSecond
          && key.equals( that.key );
    }
    return false;
} #\urlbox#
\end{minted}

Alternatively, the following listing shows another example of the \thisp{} pattern.
But in this case,
a \code{getClass} comparison is used to guard for the same type as the receiver in line 4 (\variant{GetClass} variant).

%https://lgtm.com/projects/g/neo4j/neo4j/snapshot/dist-15760049-1519892555006/files/community/bolt/src/main/java/org/neo4j/bolt/v1/messaging/infrastructure/ValuePath.java?sort=name&dir=ASC&mode=heatmap&excluded=false#L278
\def\urlvar{http://bit.ly/neo4j_neo4j_2vKP0MW}
\begin{minted}[highlightlines=7]{java}
@Override
public boolean equals( Object o ) {
    if ( this == o ) return true;
    if ( o == null || getClass() != o.getClass() )
        return false;

    ValuePath that = (ValuePath) o;
    return nodes.equals(that.nodes) &&
        relationships.equals(that.relationships);
} #\urlbox#
\end{minted}

In some situations, the type cast is not the same as the enclosing class.
Instead, the target type of the cast is the super class or a super interface of the enclosing class (\variant{InstanceOfSupertype} variant).
The following example shows this scenario.
The cast is performed in the \code{WildcardTypeImpl} enclosing class,
but the target type is \code{java.lang.reflect.WildcardType}.

%https://lgtm.com/projects/g/elastic/elasticsearch/snapshot/dist-1916470085-1524814812150/files/server/src/main/java/org/elasticsearch/common/inject/internal/MoreTypes.java?sort=name&dir=ASC&mode=heatmap#L606
\def\urlvar{http://bit.ly/elastic_elasticsearch_2GHyPp5}
\begin{minted}[highlightlines=5]{java}
public static class WildcardTypeImpl implements WildcardType, CompositeType {
    @Override
    public boolean equals(Object other) {
        return other instanceof WildcardType
                && MoreTypes.equals(this, (WildcardType) other);
    }
} #\urlbox#
\end{minted}

Similar to the previous example, 
the \variant{AutoValue} variant casts the \code{equals} parameter to a super class of the enclosing class.
However, this happens when the Google AutoValue library%
\footnote{\url{https://github.com/google/auto/tree/master/value}}
is used.
AutoValue is a code generator for value classes.

%https://lgtm.com/projects/g/square/sqlbrite/snapshot/3a9916985485ba5922097fe59a18230500f02df4/files/sample/build/generated/source/apt/debug/com/example/sqlbrite/todo/ui/$AutoValue_ListsItem.java?sort=name&dir=ASC&mode=heatmap&showExcluded=false#L52
\def\urlvar{http://bit.ly/square_sqlbrite_2HmHMYE}
\begin{minted}[highlightlines=13]{java}
@AutoValue
abstract class ListsItem implements Parcelable {
    // [...]
}

abstract class $AutoValue_ListsItem extends ListsItem {
    @Override
    public boolean equals(Object o) {
      if (o == this) {
        return true;
      }
      if (o instanceof ListsItem) {
        ListsItem that = (ListsItem) o;
        return (this.id == that.id())
             && (this.name.equals(that.name()))
             && (this.itemCount == that.itemCount());
      }
      return false;
    }
} #\urlbox#
\end{minted}

The following snippet shows a non-trivial implementation of \code{equals}.
The enclosing class of the \code{equals} method is \code{CapReq} (line 1).
However, the cast instance (line 13) is not against the enclosing class,
it is against to the \code{Requirement} class (\variant{InstanceOfSwitch} variant).
Note that the cast using the enclosing class as target type is in line 9.

%https://lgtm.com/projects/g/bndtools/bnd/snapshot/dist-930051-1524814812150/files/biz.aQute.bndlib/src/aQute/bnd/osgi/resource/CapReq.java?sort=name&dir=ASC&mode=heatmap#L73
\def\urlvar{http://bit.ly/bndtools_bnd_2SM5pOw}
\begin{minted}[highlightlines=13]{java}
class CapReq {
    @Override
    public boolean equals(Object obj) {
        if (this == obj)
                return true;
        if (obj == null)
                return false;
        if (obj instanceof CapReq)
                return equalsNative((CapReq) obj);
        if ((mode == MODE.Capability) && (obj instanceof Capability))
                return equalsCap((Capability) obj);
        if ((mode == MODE.Requirement) && (obj instanceof Requirement))
                return equalsReq((Requirement) obj);
        return false;
    }
} #\urlbox#
\end{minted}


\issues{}
The pattern for an \code{equals} method implementation is well-known.
Most \code{equals} methods in our sample are implemented with the same boilerplate structure:
that is, first checking if the parameter is another reference to \code{this},
then checking if the argument is not null,
and finally, checking if the argument is of the right class
(with either an \code{instanceof} test or a \code{getClass} comparison).
Once all checks are performed, a cast follows, and a field-by-field comparison is made.

To avoid this boilerplate, other languages bake in deep equality comparisons,
at least for some types (\eg, \scala{} case classes),
or provide mechanisms to generate the boilerplate code (\eg, \code{deriving Eq}
in \haskell{} or \code{\#[derive(Eq)]} in \rust{}).
\cite{vaziriDeclarativeObjectIdentity2007} propose a declarative approach to avoid boilerplate code when implementing
both the \code{equals} and \code{hashCode} methods.
They manually analyzed several applications, and found there are many issues while implementing \code{equals()} and \code{hashCode()} methods.
It would be interesting to check whether these issues happen in real application code.

There is an exploratory document%
\footnote{\url{http://cr.openjdk.java.net/\~briangoetz/amber/datum.html}}
by Brian Goetz---\java{} Language Architect---addressing these issues from a more general perspective.
It is definitely a starting point towards improving the \java{} language.

This pattern can be seen as a special instance of the \nameref{pat:Typecase} pattern when the guard is an \code{instanceof} test or a \code{getClass} comparison.

\end{pattern}