\begin{pattern}{NewDynamicInstance}
In the \thisp{} pattern, a new 
object or array is created by means of reflection.
The type of the object being created is determined at run time,
and the new object is cast to some statically known supertype of the run time type.

The \code{newInstance} method family declared in the \code{Class},
% \footnote{\url{https://docs.oracle.com/javase/8/docs/api/java/lang/Class.html\#newInstance--}}
\code{Array}
% \footnote{\url{https://docs.oracle.com/javase/8/docs/api/java/lang/reflect/Array.html\#newInstance-java.lang.Class-int-}}\(^{,}\)
% \footnote{\url{https://docs.oracle.com/javase/8/docs/api/java/lang/reflect/Array.html\#newInstance-java.lang.Class-int...-}}
and \code{Constructor}
% \footnote{\url{https://docs.oracle.com/javase/8/docs/api/java/lang/reflect/Constructor.html\#newInstance-java.lang.Object...-}}
classes creates an object or array dynamically by means of reflection, \ie,
the type of object being created is not known at compile-time.
This pattern consists of casting the result of these methods to the appropriate target type.

\instances{}
The following example%
\def\urlvar{http://bit.ly/apache_hadoop_2HC3IPg}
shows a cast of the result of the \code{Class.newInstance()}
method.

%https://lgtm.com/projects/g/apache/hadoop/snapshot/6bedbef6c5f2d937a6cbc268300ce2a39609d06c/files/hadoop-hdfs-project/hadoop-hdfs/src/main/java/org/apache/hadoop/hdfs/server/namenode/FSNamesystem.java?sort=name\&dir=ASC\&mode=heatmap\&showExcluded=false#L1039
\begin{minted}[highlightlines=1,linenos=false]{java}
logger = (AuditLogger) Class.forName(className).newInstance();
#\urlbox#
\end{minted}

The following example%
\def\urlvar{http://bit.ly/neo4j_neo4j_2Hp5Hqc}
shows how to dynamically create an array, using the \texttt{Array} class.

%https://lgtm.com/projects/g/neo4j/neo4j/snapshot/27aaa67633e4d26446e38125d04fbbd27f938b75/files/community/collections/src/main/java/org/neo4j/helpers/collection/Iterables.java?sort=name\&dir=ASC\&mode=heatmap\&showExcluded=false#L403
\begin{minted}[highlightlines=1]{java}
return list.toArray( (T[]) Array.newInstance( componentType, list.size()));
#\urlbox#
\end{minted}

Whenever a constructor other than the default constructor is needed,
the \code{newInstance} method declared in the \code{Constructor} class
should be used to select the appropriate constructor,
as shown in the following example.%
\def\urlvar{http://bit.ly/gradle_gradle_2HsUgOo}

%https://lgtm.com/projects/g/gradle/gradle/snapshot/209c3175e75af6ac30cb66c02eda15b0f8b6a616/files/subprojects/internal-integ-testing/src/main/groovy/org/gradle/integtests/fixtures/executer/OutputScrapingExecutionFailure.java?sort=name\&dir=ASC\&mode=heatmap\&showExcluded=false#L174
\begin{minted}[highlightlines=1]{java}
return (Exception) Class
                       .forName(className)
                       .getConstructor(String.class)
                       .newInstance(message);
#\urlbox#
\end{minted}

The following example%
\def\urlvar{http://bit.ly/alibaba_LuaViewSDK_2HC33xg}
shows a guarded instance of the \thisp{} pattern.
This seems rather unusual, as this pattern is not guarded.

%https://lgtm.com/projects/g/alibaba/LuaViewSDK/snapshot/dist-2037250419-1524814812150/files/Android/LuaViewSDK/src/com/taobao/luaview/global/LuaViewManager.java?sort=name&dir=ASC&mode=heatmap#L373
\begin{minted}[highlightlines=6]{java}
private static List<String> getMapperMethodNames(final Class clazz) {
    try {
        if (clazz != null) {
            Object obj = clazz.newInstance();
            if (obj instanceof BaseMethodMapper) {
                return ((BaseMethodMapper) obj).getAllFunctionNames();
            }
        }
    } catch (Exception e) {
        e.printStackTrace();
    }
    return null;
} #\urlbox#
\end{minted}

There are cases when the cast is not directly applied to the result of the \code{newInstance} method.
The following snippet shows such a case.%
\def\urlvar{http://bit.ly/pac4j_pac4j_2HJtXUn}
The cast is used to convert from \code{Class<?>} to \code{Class<ConfigFactory>} (line 4).
The invocation to the \code{newInstance} method then does not need a direct cast (line 8) given the definition of the \code{clazz} variable (line 2).
Nevertheless, the cast is unchecked, and a \code{checkcast} instruction is going to be emitted anyway for the result of the \code{newInstance} invocation.

%https://lgtm.com/projects/g/pac4j/pac4j/snapshot/dist-4840350-1524814812150/files/pac4j-core/src/main/java/org/pac4j/core/config/ConfigBuilder.java?sort=name&dir=ASC&mode=heatmap#L25
\begin{minted}[highlightlines=4]{java}
ClassLoader tccl = Thread.currentThread().getContextClassLoader();
final Class<ConfigFactory> clazz;
if (tccl == null) {
    clazz = (Class<ConfigFactory>) Class.forName(factoryName);
} else {
    clazz = (Class<ConfigFactory>) Class.forName(factoryName, true, tccl);
}
final ConfigFactory factory = clazz.newInstance();
#\urlbox#
\end{minted}



\discussion{}
The cast here is needed because of the dynamic nature of reflection.
This pattern is usually unguarded; that is,
the programmer knows what target type is being created.

Generics could be used to avoid the cast on \code{newInstance}, assuming the \code{Class} instance is 
not a raw type or a \code{Class<?>}.
However, the usual API for getting a class instance \code{Class.forName}
returns such a type.
Indeed, the following two snippets:

\begin{minted}[highlightlines=2,linenos=false]{java}
Class<?> c = Class.forName("java.lang.String");
String pf = (String) c.newInstance();
\end{minted}

\begin{minted}[highlightlines=1,linenos=false]{java}
Class<String> c = (Class<String>) Class.forName("java.lang.String");
String pf = c.newInstance();
\end{minted}

compile to the same bytecode below.

\begin{lstlisting}[style=bytecode]
ldc           #24    // String java.lang.String
invokestatic  #26    // Method java/lang/Class.forName
astore_1
aload_1
invokevirtual #32    // Method java/lang/Class.newInstance
checkcast     #36    // class java/lang/String
\end{lstlisting}


In the first case, the cast is on the \code{newInstance}, an instance of the
\thisp{} pattern. In the second case, the cast is on the call to
\code{Class.forName}, an instance of the \nameref{pat:Factory} pattern.

This pattern is related to \nameref{pat:Deserialization},
since both create an object dynamically.
It is also related to \nameref{pat:ReflectiveAccessibility},
where both retrieve objects by reflection.

\end{pattern}