\begin{pattern}{SoleSubclassImplementation}
The \thisp{} occurs when an interface or abstract class has only one implementing subclass.
Casting the interface to this class must succeed because it cannot possibly be an instance of another class.

\instances{}
In the following example the \code{jobId}
variable is cast to the sole implementation (\code{JobIdImpl}).

%https://lgtm.com/projects/g/ow2-proactive/scheduling/snapshot/dist-6910096-1524814812150/files/scheduler/scheduler-api/src/main/java/org/ow2/proactive/scheduler/job/JobIdImpl.java?sort=name&dir=ASC&mode=heatmap#L118
\def\urlvar{http://bit.ly/ow2_proactive_scheduling_2Ulcjfs}
\begin{minted}[highlightlines=1,linenos=false]{java}
return Longs.compare(id, ((JobIdImpl) jobId).id);
#\urlbox#
\end{minted}

Similar to the previous example,
the variable \code{user} is cast to the known implementation (\code{UserImpl}).

%https://lgtm.com/projects/g/Javacord/Javacord/snapshot/dist-1791982360-1524814812150/files/src/main/java/de/btobastian/javacord/entities/message/impl/ImplMessage.java?sort=name&dir=ASC&mode=heatmap#L681
\def\urlvar{http://bit.ly/Javacord_Javacord_2GwGjuV}
\begin{minted}[highlightlines=2]{java}
for (User user : api.getUsers()) {
    if (channelId.equals(((ImplUser) user).getUserChannelId())) {
        return user;
    }
} #\urlbox#
\end{minted}


\detection{}
The following query returns all casts such that the type---class or interface---of the expression being cast has only one subtype.
The \emph{transitive closure} symbol \code{+} indicates that \code{getASubtype} may be followed one or more times.

\begin{listing}
\begin{minted}{\qllexer}
class SoleSubclassImplementation extends #\qlref{Cast}# { #\qlbox#
	SoleSubclassImplementation() {
		count(RefType rt | 
				rt = getExpr().getType() and rt.fromSource() | 
				rt.getASubtype+() ) = 1
	}
}
\end{minted}
\caption{Detection of the \thisp{} pattern.}
\label{lst:ql:SoleSubclassImplementationCast}
\end{listing}


\issues{}
This pattern occurs when there is high cohesion between super and subclass.
In some cases, the cast instance appears in a generated class.
This mechanism allows the developer to extend this class to add custom code.
Therefore, this high cohesion is acceptable.
The developer assumes that there is no other implementation of the base class,
otherwise the cast instance fails.

\end{pattern}
