\begin{pattern}{ParserStack}
The \thisp{} pattern consists of multiple cases,
dispatched depending on some application-specific control state,
with casts of the top elements of stack-like collection in each case.
An application invariant ensures that if the application is in a given 
state then the top elements of the stack should be of known run-time types.

\instances{}
The following example,
shows a cast whose value is on top of a stack (line 2).
In this case,
the code is transforming a parse tree into an abstract syntax tree.
The casts in the switch case are guarded by the parse tree node type and
its arity. 

%https://lgtm.com/projects/g/fabioz/Pydev/snapshot/dist-20832102-1524814812150/files/plugins/org.python.pydev.parser/src/org/python/pydev/parser/grammar27/TreeBuilder27.java?sort=name&dir=ASC&mode=heatmap#L231
\def\urlvar{http://bit.ly/fabioz_Pydev_2HF6nrF}
\begin{minted}[highlightlines=2]{java}
case JJTASSERT_STMT:
    exprType msg = arity == 2 ? ((exprType) stack.popNode()) : null;
    test = (exprType) stack.popNode();
    return new Assert(test, msg);
#\urlbox#
\end{minted}

Similar to the previous example,
in this case a guarded cast is performed on a stack of grammar symbols.
The code was generated using an LR parser generator.
The guard ensures that the parser has already matched a given prefix of the
input and so the top of the stack should contain the expected symbols.

%https://lgtm.com/projects/g/Sable/soot/snapshot/dist-1791462132-1524814812150/files/src/main/generated/jastadd/soot/JastAddJ/JastAddJavaParser.java?sort=name&dir=ASC&mode=heatmap#L1036
\def\urlvar{http://bit.ly/Sable_soot_2MZLZ3m}
\begin{minted}[highlightlines=4]{java}
case 40: // qualified_name_decl = name_decl.n DOT.DOT IDENTIFIER.i
{
    final Symbol _symbol_n = _symbols[offset + 1];
    final IdUse n = (IdUse) _symbol_n.value;
    final Symbol DOT = _symbols[offset + 2];
    final Symbol i = _symbols[offset + 3];
    return new IdUse(n.getID() + "." + ((String)i.value));
} #\urlbox#
\end{minted}


\detection{}
To manually detect this pattern,
we look for methods that pop up an element from a stack,
and then cast of it.
Automatic detection for this pattern becomes impractical,
since a query would need to detect such a method,
and when a class is implementing a stack-like structure.


\issues{}
In our sample,
this pattern is always seen when implementing grammar-related operations,
such as parsers or interpreters.
In some situations, similar to the \nameref{pat:Stash} pattern,
this pattern could be replaced with a strongly typed heterogenous collection~\citep{kiselyovStronglyTypedHeterogeneous2004}.

Similar to \nameref{pat:Typecase},
multiple cases are evaluated with casts to different types,
depending on application-specific guards.
However, unlike \nameref{pat:Typecase}, 
the success of the casts is ensured not by a type-tag-like value,
but by application-specific state
(\eg{}, the current parser state or the state of an evaluator)
and proper use of the stack.

\end{pattern}
