\begin{pattern}{Literal}
Cast a numeric or character literal or constant --- defined as \code{static final} ---
to a primitive type.

\instances{}
Casting a literal to initialize an array is shown in the following listing.%
\footnote{\url{http://bit.ly/2Fz3fuN}}
This kind of cast is seen often.
Note that actually this kind is redundant.
It is just there because it was a generated cast.

% https://lgtm.com/projects/g/robovm/robovm/snapshot/dist-39650108-1524814812150/files/rt/external/bouncycastle/bcprov/src/main/java/org/bouncycastle/crypto/engines/RC2Engine.java?sort=name&dir=ASC&mode=heatmap#L23
\begin{minted}[highlightlines=5]{java}
private static byte[] piTable =
{
    (byte)0xd9, (byte)0x78, (byte)0xf9, (byte)0xc4, (byte)0x19, (byte)0xdd,
    (byte)0xb5, (byte)0xed, (byte)0x28, (byte)0xe9, (byte)0xfd, (byte)0x79,
    (byte)0x4a, (byte)0xa0, (byte)0xd8, (byte)0x9d, 
\end{minted}

The following listing shows an example of the \pname{Literal} pattern.%
\footnote{\url{http://bit.ly/2FuEqkC}}
% https://lgtm.com/projects/g/kaitoy/pcap4j/snapshot/dist-29675155-1524814812150/files/pcap4j-core/src/main/java/org/pcap4j/packet/namednumber/TcpPort.java?sort=name&dir=ASC&mode=heatmap#L2329
% \begin{listing}
\begin{minted}[highlightlines=2]{java}
public static final TcpPort POWERBURST =
    new TcpPort((short)485, "Air Soft Power Burst");
\end{minted}
% \caption{Literal example (\url{http://bit.ly/2FuEqkC} )}
% \end{listing}

Another example using a constant instead of a literal:%
\footnote{\url{http://bit.ly/2FuBp3M}}

% https://lgtm.com/projects/g/jline/jline2/snapshot/dist-4850314-1524814812150/files/src/test/java/jline/console/ConsoleReaderTest.java?sort=name&dir=ASC&mode=heatmap#L164
\begin{minted}[highlightlines=6]{java}
char[] characters = new char[]{
    'S', 's',
    (char) NUMPAD_KEY_INDICATOR.code,
    (char) LEFT_ARROW_KEY.code,
    (char) NUMPAD_KEY_INDICATOR.code,
    (char) DELETE_KEY.code, '\r', 'n'
};
\end{minted}

\detection{}
This pattern looks for casting a literal or constant to a primitive type.

\discussion{}
\java{} already provides several types of numeric literals:
\code{int}, \code{long}, \code{float}.

This pattern is related with the \nameref{pat:Prim} pattern.
Also it is related with the \nameref{pat:SelectOverload}
because it does not permit for narrowing conversions.%
\footnote{\url{https://docs.oracle.com/javase/specs/jls/se8/html/jls-5.html\#jls-5.3}}
Related to \nameref{pat:Redundant} because of cast automatically generated.
\end{pattern}