\begin{pattern}{PatternMatching}
This pattern is composed of a guard (\code{instanceof}) followed by a cast to known subtypes of the static type.
Often there is just one case and the default case, \ie, \code{instanceof} fails, does a no-op or reports an error.
Another common approach is to have several cases,
usually one \emph{per} subtype.

\instances{}
The following listing shows an example of the \thisp{} pattern.%
\footnote{\url{http://bit.ly/OpenMods_OpenBlocks_2FzYYHq}}
%
\done{Nate: Highlightlines color difficult to read}
%
In this example, there is only a single case (line 3).
After the cast is done, a specific operation is performed on the cast type,
\ie, invoking the \code{checkBlock} method.

%https://lgtm.com/projects/g/OpenMods/OpenBlocks/snapshot/dist-2040060754-1524814812150/files/build/sources/java/openblocks/common/tileentity/TileEntityImaginary.java?sort=name&dir=ASC&mode=heatmap#L268
\begin{minted}[highlightlines=3]{java}
Item item = helmet.getItem();
if (item instanceof ItemImaginationGlasses)
	return ((ItemImaginationGlasses)item).checkBlock(what, helmet, this);
\end{minted}

In the next case,%
\footnote{\url{http://bit.ly/PenguinSquad_Enchiridion_2HnNwB7}}
more than one type test are performed.
Each cast corresponds to a type test.

%https://lgtm.com/projects/g/PenguinSquad/Enchiridion/snapshot/dist-19218583-1524814812150/files/build/sources/main/java/joshie/enchiridion/helpers/StackHelper.java?sort=name&dir=ASC&mode=heatmap#L27
\begin{minted}[highlightlines=7]{java}
public static String getStringFromObject(Object object) {
	if (object instanceof Item) {
		return getStringFromStack(new ItemStack((Item) object));
	} else if (object instanceof Block) {
		return getStringFromStack(new ItemStack((Block) object));
	} else if (object instanceof ItemStack) {
		return getStringFromStack((ItemStack) object);
	} else if (object instanceof String) {
		return (String) object;
	} else if (object instanceof List) {
		return getStringFromStack((ItemStack) ((List) object).get(0));
	} else return "";
}
\end{minted}

There are situations when the type test is applied to more than one variable.
In the following example%
\footnote{\url{http://bit.ly/bbossgroups_bboss_2FDN9Rd}}
a double type test is performed on the parameters \code{o1} and \code{o2}.

%https://lgtm.com/projects/g/bbossgroups/bboss/snapshot/dist-2025970729-1524814812150/files/bboss-util/src/org/frameworkset/util/ObjectUtils.java?sort=name&dir=ASC&mode=heatmap#L228
\begin{minted}[highlightlines=25]{java}
public static boolean nullSafeEquals(Object o1, Object o2) {
	if (o1 == o2) {
		return true;
	}
	if (o1 == null || o2 == null) {
		return false;
	}
	if (o1.equals(o2)) {
		return true;
	}
	if (o1.getClass().isArray() && o2.getClass().isArray()) {
		if (o1 instanceof Object[] && o2 instanceof Object[]) {
			return Arrays.equals((Object[]) o1, (Object[]) o2);
		}
		if (o1 instanceof boolean[] && o2 instanceof boolean[]) {
			return Arrays.equals((boolean[]) o1, (boolean[]) o2);
		}
		if (o1 instanceof byte[] && o2 instanceof byte[]) {
			return Arrays.equals((byte[]) o1, (byte[]) o2);
		}
		if (o1 instanceof char[] && o2 instanceof char[]) {
			return Arrays.equals((char[]) o1, (char[]) o2);
		}
		if (o1 instanceof double[] && o2 instanceof double[]) {
			return Arrays.equals((double[]) o1, (double[]) o2);
		}
		if (o1 instanceof float[] && o2 instanceof float[]) {
			return Arrays.equals((float[]) o1, (float[]) o2);
		}
		if (o1 instanceof int[] && o2 instanceof int[]) {
			return Arrays.equals((int[]) o1, (int[]) o2);
		}
		if (o1 instanceof long[] && o2 instanceof long[]) {
			return Arrays.equals((long[]) o1, (long[]) o2);
		}
		if (o1 instanceof short[] && o2 instanceof short[]) {
			return Arrays.equals((short[]) o1, (short[]) o2);
		}
	}
	return false;
}
\end{minted}

Another common scenario%
\footnote{\url{http://bit.ly/codefollower_Tomcat-Research_2SGDUG5}}
is when several cases are used to re-\code{throw} an exception of the right type, as shown below.

%https://lgtm.com/projects/g/codefollower/Tomcat-Research/snapshot/dist-13734061-1524814812150/files/java/org/apache/naming/factory/DataSourceLinkFactory.java?sort=name&dir=ASC&mode=heatmap#L85
\begin{minted}[highlightlines=10]{java}
try {
	// [...]
}catch (Exception x) {
	if (x instanceof InvocationTargetException) {
		Throwable cause = x.getCause();
		if (cause instanceof ThreadDeath) {
			throw (ThreadDeath) cause;
		}
		if (cause instanceof VirtualMachineError) {
			throw (VirtualMachineError) cause;
		}
		if (cause instanceof Exception) {
			x = (Exception) cause;
		}
	}
	// [...]
}
\end{minted}

The next example%
\footnote{\url{http://bit.ly/kiegroup_jbpm_2ENCL8a}}
shows how \thisp{} can be used to filter elements by type within a stream.
The cast instance is applied to stream operations (line 3) over the \code{caseAssignments} collection.
The filter operation by type is performed in line 4.

%https://lgtm.com/projects/g/kiegroup/jbpm/snapshot/dist-1810050-1524814812150/files/jbpm-case-mgmt/jbpm-case-mgmt-impl/src/main/java/org/jbpm/casemgmt/impl/wih/StartCaseWorkItemHandler.java?sort=name&dir=ASC&mode=heatmap#L212
\begin{minted}[highlightlines=3]{java}
Collection<OrganizationalEntity> caseAssignments = 
		((CaseAssignment)caseFile).getAssignments(name);
user = (User) caseAssignments.stream()
		.filter(oe -> oe instanceof User)
		.findFirst()
		.orElseThrow(() -> new IllegalArgumentException());

public interface User extends OrganizationalEntity {}
public interface OrganizationalEntity {}
\end{minted}


\detection{}
To detect this pattern,
a cast needs to be guarded by an \code{instanceof} expression.
The operand expression to both the cast and the \code{instanceof} needs to be the same.
In the case the operand expression is a variable, we also check that there is no assignment between the \code{instanceof} guard and the cast to that variable.

In some situations the operand expression is a method invocation.
%
\discuss{Nate: Avoid "we"}
%
We make sure that the value returned is the same for both the \code{instanceof} and the cast, \ie, it is a pure method.

\discussion{}
The \thisp{} pattern consists of testing the runtime type of a variable against several related types.
It is a technique that allows a developer to take different actions according to the runtime type of an object.
Depending on the --- runtime --- type of an object, different cases, usually one for each type will follow.

The \thisp{} pattern can be seen as an \adhoc{} alternative to pattern matching as a language construct~\citep{lavilleLazyPatternMatching1987}.
This construct can be seen in several other languages, \eg, \scala{}, \csharp{}, and \haskell{}.
For instance, in \scala{} the pattern matching construct is achieved using the \code{match} keyword.
In this example,%
\footnote{Adapted from \url{https://docs.scala-lang.org/tour/pattern-matching.html}}
a different action is taken according to the runtime type of the parameter \code{notification} (line 9).

\begin{minted}[highlightlines=10]{scala}
abstract class Notification
case class Email(sender: String, title: String, body: String)
	extends Notification
case class SMS(caller: String, message: String)
	extends Notification
case class VoiceRecording(contactName: String, link: String)
	extends Notification

def showNotification(notification: Notification): String = {
	notification match {
		case Email(email, title, _) =>
		s"You got an email from $email with title: $title"
		case SMS(number, message) =>
		s"You got an SMS from $number! Message: $message"
		case VoiceRecording(name, link) =>
		s"Voice Recording from $name! Click the link: $link"
	}
	}
	val someSms = SMS("12345", "Are you there?")
	val someVoiceRecording = VoiceRecording("Tom", "voicerecording.org/id/123")
	
	// prints You got an SMS from 12345! Message: Are you there?
	println(showNotification(someSms))
	
	// Voice Recording from Tom! Click the link: voicerecording.org/id/123	
	println(showNotification(someVoiceRecording))
\end{minted}

As a workaround, alternatives to the \thisp{} pattern can be the visitor pattern or polymorphism.
But in some cases, the chain of \code{instanceof}s is of boxed types.
Thus no polymorphism can be used.
There is an ongoing proposal%
\footnote{\url{http://openjdk.java.net/jeps/305}}$^{,}$%
\footnote{\url{https://cr.openjdk.java.net/~briangoetz/amber/pattern-match.html}}
to add pattern matching to the \java{} language.
The proposal explores how to change the \code{instanceof} operator in order to support pattern matching.
%
\done{Java 12 will (maybe) add matching.}
%
\java{} 12 already extends the \code{switch} statement to be used as either a statement or an expression.%
\footnote{\url{https://openjdk.java.net/jeps/325}}
This enhancement aims to ease the transition to a \code{switch} expression that supports pattern matching.

\related{}
The \nameref{pat:Equals} pattern can be seen as a special case of the \thisp{} pattern,
because often the parameter to the \code{equals} method is guarded by an \code{instanceof} expression.
\end{pattern}
