\section{Research Questions}

The examples in the previous section indicate that type casts represent a source of errors for developers.
Therefore we want to understand why developers need to use type casts.
To this end, we propose to answer the following question:
\emph{How and when do developers need to escape the type system?}
%
\done{Nate: Should be specific to casts?}
%
The cast operator in \java{} provides the means to view a reference as a different type to how it was declared.
%
\done{Nate: Define upcasts}
%
Upcasts are done automatically by the compiler.
Nevertheless, as we shall see later, in some situations a developer is forced to insert upcasts.
In the case of downcasts,
a check is inserted at run-time to verify that the conversion is sound,
%
\done{Nate: Remove "thus escaping the static type system."}
%
thus turning a static type test into a dynamic test.

Therefore, we believe we should care about how the casting operations are used in the wild.
Specifically, we want to answer the following research questions:

\begin{enumerate}[label=$RQ/C\arabic*:$,ref=$RQ/C\arabic*$,leftmargin=3.4\parindent]
\item\label{enum:rq1}{\bf \crqA}
We want to understand to what extent application code actually uses cast operations.
\item\label{enum:rq2}{\bf \crqB}
If casts are actually used in application code, we want to know how and why developers need to escape the type system.
\item\label{enum:rq3}{\bf \crqC}
In addition to understand how and when casts are used, we want to measure how often developers need to resort to certain idioms to solve a particular problem.
\end{enumerate}

To answer the above questions, we need to determine whether and how cast operations are actually used in real-world \java{} applications.
In Section~\ref{sec:casts:stats} we first give an estimation of how often the cast operator is used in real-world applications to answer~\ref{casts:rq1}.
In Section~\ref{sec:casts:methodology} we introduce the methodology we have used to devise cast usage patterns.
Then, Section~\ref{sec:casts:patterns} presents the cast usage patterns we have devised to answer both~\ref{casts:rq2} and \ref{casts:rq3}.
